---
title: "607 - Chess Tournament Project"
author: "Karim Hammoud"
date: "`r Sys.Date()`"
output:
  html_document:
    highlight: pygments
    theme: cerulean
    toc: yes
    toc_float: yes
  pdf_document: default
  openintro::lab_report: default
editor_options:
  chunk_output_type: console
---

```{r message=FALSE, echo=FALSE}
knitr::opts_chunk$set(eval = TRUE, results = TRUE, fig.show = "show", message = TRUE)
```

## Overview
In this project, you’re given a text file with chess tournament results where the information has some structure. Your job is to create an R Markdown file that generates a .CSV file (that could for example be imported into a SQL database) with the following information for all of the players:

Player’s Name, Player’s State, Total Number of Points, Player’s Pre-Rating, and Average Pre Chess Rating of Opponents

For the first player, the information would be:

Gary Hua, ON, 6.0, 1794, 1605

1605 was calculated by using the pre-tournament opponents’ ratings of 1436, 1563, 1600, 1610, 1649, 1663, 1716, and dividing by the total number of games played.

The chess rating system (invented by a Minnesota statistician named Arpad Elo) has been used in many other contexts, including assessing relative strength of employment candidates by human resource departments.

**[Github link here](https://github.com/akarimhammoud/CUNY-SPS/tree/master/607-Data-Acquisition-and-Management-CUNY-SPS-Fall2020/Week4-Project1)**

**[Rpubs link here](https://rpubs.com/karim7mod/662977)**

## Import the required libraries

```{r message=FALSE}
library(tidyverse)
library(openintro)
library(stringr)
```

## Read the text file

```{r}
url <- "https://raw.githubusercontent.com/akarimhammoud/CUNY-SPS/master/607-Data-Acquisition-and-Management-CUNY-SPS-Fall2020/Week4-Project1/tournamentinfo.txt"

tournamentinfo <- read.csv(paste0(url), header=F)
head (tournamentinfo)
tail(tournamentinfo)
```

## Data wrangling 

Taking out the first four rows

```{r}
tournamentinfo <- tournamentinfo[-c(1:4),]
head(tournamentinfo)
```

Checking the remaining rows

```{r}
length(tournamentinfo)
```

## Pulling the first and second rows of each three rows.

```{r}
first_row <- tournamentinfo[seq(1, length(tournamentinfo), 3)]
head(first_row,2)
second_row <- tournamentinfo[seq(2, length(tournamentinfo), 3)]
head(second_row,2)
```

## Using regular expression to extract the Data.

```{r}
#pair number
number <- as.integer(str_extract(first_row,'\\d+'))
number
#player's name
player_name <- str_trim(str_extract(first_row, '(\\w+\\s){2,3}'))
#player's state
player_state <- str_extract(second_row, "\\w+")
#points
player_points <- as.numeric(str_extract(first_row, '\\d+\\.\\d+'))
head(player_points)
# the rating
player_rating <- as.integer(str_extract(str_extract(second_row, '[^\\d]\\d{3,4}[^\\d]'), '\\d+'))
head(player_rating)
# the opponents
opponents <- str_extract_all(str_extract_all(first_row, "\\d+\\|"), "\\d+")
head(opponents)
#count the result
won <- str_count(first_row,'\\Q|W \\E')
lost <- str_count(first_row, '\\Q|L, \\E')
draw <- str_count(first_row, '\\Q|D \\E')
```

## Calculate the mean rating

```{r}
mean_rating <- length(first_row)

for (i in 1:length(first_row)) { 
  mean_rating[i] <- round(mean(player_rating[as.numeric(unlist(opponents[number[i]]))]), digits = 0) 
}
```

## The final data frame

```{r}
final_data <- data.frame(player_name, player_state, player_points, player_rating, mean_rating)
head(final_data)
```

## Change the heading names

```{r}
colnames(final_data) <- c("Name", "State", "Points", "Rating", "Average Rating")
head(final_data)
tail(final_data)
```

## Create the CSV file in the general folder in Mac

```{r}
write.csv(final_data, file = "/Users/karimh/Documents/R practices/Project1_607.csv")
```
