---
title: "Chapter 2 - Summarizing Data"
author: "Karim Hammoud"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

### Exercise 1

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Stats scores**. (2.33, p. 78) Below are the final exam scores of twenty introductory statistics students. 

57, 66, 69, 71, 72, 73, 74, 77, 78, 78, 79, 79, 81, 81, 82, 83, 83, 88, 89, 94

Create a box plot of the distribution of these scores. The five number summary provided below may be useful.


\begin{center}
\renewcommand\arraystretch{1.5}
\begin{tabular}{ccccc}
Min & Q1    & Q2 (Median)   & Q3    & Max \\
\hline
57  & 72.5  & 78.5          & 82.5  & 94 \\
\end{tabular}
\end{center}

```{r, echo = FALSE}
scores <- c(57, 66, 69, 71, 72, 73, 74, 77, 78, 78, 79, 79, 81, 81, 82, 83, 83, 88, 89, 94)
```

* The answer: boxplot for the distribution of the scores
```{r blot}
boxplot(scores, main = 'Final exam scores of twenty introductory statistics students', ylab = 'scores', ylim = c(55, 100))
```

* quantile for the boxplot
```{r quantile}
quantile(scores, probs = c(0,.25,.5,.75,1))
```

\clearpage

### Exercise 2

**Mix-and-match**. (2.10, p. 57) Describe the distribution in the histograms below and match them to the box plots.

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=3}
library(openintro)
set.seed(7365)
sym = rnorm(1000, mean = 60, sd = 3)
uni = runif(1000, min = 0, max = 100)
rs = rgamma(1000, shape = 3, rate = 2)
par(mar=c(4, 3.6, 0, 0), las = 1, mgp = c(2.7, 0.7, 0), 
    mfrow = c(1,6), 
    cex.lab = 1.5, cex.axis = 1.5)
histPlot(sym, xlab = "(a)", ylab = "", col = COL[1], axes = FALSE)
axis(1, seq(50,70,10))
histPlot(uni, xlab = "(b)", ylab = "", col = COL[1], axes = FALSE)
axis(1, seq(0,100,50))
histPlot(rs, xlab = "(c)", ylab = "", col = COL[1], axes = FALSE)
axis(1, seq(0,6,2))
boxPlot(rs, xlab = "(1)", ylab = "", col = COL[1,3])
boxPlot(sym, xlab = "(2)", ylab = "", col = COL[1,3])
boxPlot(uni, xlab = "(3)", ylab = "", col = COL[1,3])
```

* The answer: The data is symmetric distribution, histograms (a) shows the data distribution between 50-70 and the majority are in the 60s, this could match with boxplot (2).

* The data is neither skewed or syemmetic and has no distribution, histograms (b) shows the data distribution between 0-100 and the data is spread between these values, this could match with boxplot (3).

* The data is right skewed distribution, histograms (c) shows the data distribution between 0-6 and the majority are between 1-2, this could match with boxplot (1).

\clearpage

### Exercise 3

**Distributions and appropriate statistics, Part II**. (2.16, p. 59) For each of the following, state whether you expect the distribution to be symmetric, right skewed, or left skewed. Also specify whether the mean or median would best represent a typical observation in the data, and whether the variability of observations would be best represented using the standard deviation or IQR. Explain your reasoning.

(a) Housing prices in a country where 25% of the houses cost below $350,000, 50% of the houses cost below $450,000, 75% of the houses cost below $1,000,000 and there are a meaningful number of houses that cost more than $6,000,000.

* The Answer: this is a right skewed distribution, median is better because the data is skewed and it is less affected by outliers, and its better to use standard deviation.

(b) Housing prices in a country where 25% of the houses cost below $300,000, 50% of the houses cost below $600,000, 75% of the houses cost below $900,000 and very few houses that cost more than $1,200,000.

* The Answer:The data is symmetric distribution, mean is better because the data is symmetric and it provides better measure of central tendency while there are no outliers, and its better to use standard deviation when there are no outliers. 

(c) Number of alcoholic drinks consumed by college students in a given week. Assume that most of these students don't drink since they are under 21 years old, and only a few drink excessively, 

* The Answer: this is a right skewed distribution when age increase the drinking between students increase and especiall after 21, median is better because the data is skewed and the data is less affected by outliers., and its better to use IQR.

(d) Annual salaries of the employees at a Fortune 500 company where only a few high level
executives earn much higher salaries than the all other employees.

* The Answer: this is a right skewed distribution as small percentage earns alot higher than other employees, median is better because the data is skewed, and its better to use IQR.

\clearpage

### Exercise 4

**Heart transplants.** (2.26, p. 76) The Stanford University Heart Transplant Study was conducted to determine whether an  experimental heart transplant program increased lifespan. Each patient entering the program was designated an official heart transplant candidate, meaning that he was gravely ill and would most likely benefit from a new heart. Some patients got a transplant and some did not. The variable *transplant* indicates which group the patients were in; patients in the treatment group got a transplant and those in the control group did not. Of the 34 patients in the control group, 30 died. Of the 69 people in the treatment group, 45 died. Another variable called *survived* was used to indicate whether or not the patient was alive at the end of the study.

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold", out.width="50%", fig.height=3}
library(openintro)
url <- "https://raw.githubusercontent.com/jbryer/DATA606Fall2020/master/course_data/os3_data/Ch%201%20Exercise%20Data/heart_transplant.csv"
heartTr <- read.csv(file= url, header=TRUE, sep = ",")
# mosaic plot -------------------------------------------------------
par(mar = c(0, 0, 0, 0), las = 1, mgp = c(2.7, 0.9, 0))
mosaicplot(transplant ~ survived, data = heartTr, 
           main = "", xlab = "", ylab = "", color = COL[1],
           cex.axis = 1.5)

# box plot ----------------------------------------------------------
par(mar = c(2, 4.8, 0, 0), las = 1, mgp = c(3.5, 0.7, 0), 
    cex.lab = 1.5, cex.axis = 1.5)
boxPlot(heartTr$survtime, fact = heartTr$transplant, 
        ylab = "Survival Time (days)", col = COL[1,2])
```

(a) Based on the mosaic plot, is survival independent of whether or not the 
patient got a transplant? Explain your reasoning.

* The Answer: No the survival is not independent for the transplant treatment as per the mosaic plot, and the ones who got treatment has much higher survival rate.

(b) What do the box plots below suggest about the efficacy (effectiveness) of the heart transplant treatment.

* The Answer:The ones who got treatment increased their survival rate therefore the treatment is effective.

(c) What proportion of patients in the treatment group and what proportion of 
patients in the control group died?

* The Answer:
```{r}
all_control <- subset(heartTr, heartTr$transplant == 'control')
dead_contrl <- subset(heartTr, heartTr$survived == 'dead' & heartTr$transplant == 'control')
nrow(dead_contrl)/nrow(all_control)

# The people who died are 88.24% of the control group.
```

```{r}
all_treat <- subset(heartTr, heartTr$transplant == 'treatment')
dead_treat <- subset(heartTr, heartTr$survived == 'dead' & heartTr$transplant == 'treatment')
nrow(dead_treat)/nrow(all_treat)

# The people who died are 65.22% of the treatment group.
```

(d) One approach for investigating whether or not the treatment is effective 
is to use a randomization technique.

i. What are the claims being tested?

* The Answer:
H0 Independent Model: The transplant treatment has no effect on survival. The observed higher survival rate was due to chance.
H1 Alternative model: The transplant treatment has effect on patient survival. The observed higher survival rate was not due to chance.

ii. The paragraph below describes the set up for such approach, if we were 
to do it without using statistical software. Fill in the blanks with a number 
or phrase, whichever is appropriate.

* The Answer:
```{r}
nrow(subset(heartTr, heartTr$survived == 'alive'))

nrow(heartTr) - nrow(subset(heartTr, heartTr$survived == 'alive'))

nrow(all_treat)

nrow(all_control)

nrow(dead_contrl)/nrow(all_control) - nrow(dead_treat)/nrow(all_treat)
```
> We write *alive* on 28 cards representing patients who were 
alive at the end of the study, and *dead* on 75 cards 
representing patients who were not. Then, we shuffle these cards and split them 
into two groups: one group of size 69 representing treatment, and 
another group of size 34 representing control. We calculate the 
difference between the proportion of *dead* cards in the treatment and 
control groups (treatment - control) and record this value. We repeat this 100 
times to build a distribution centered at 0. Lastly, we calculate 
the fraction of simulations where the simulated differences in proportions are 
23.02%. If this fraction is low, we conclude that it is unlikely to 
have observed such an outcome by chance and that the null hypothesis should 
be rejected in favor of the alternative.

iii. What do the simulation results shown below suggest about the effectiveness 
of the transplant program?

* The Answer: the differences in proportions are 23.02% which is not high, therefore the null hypothesis is reject.

\end{center}
