---
title: "607 - Web Technologies"
author: "Karim Hammoud - karimalhammoud@gmail.com"
date: "`r Sys.Date()`"
output:
  html_document:
    highlight: pygments
    theme: cerulean
    toc: yes
    toc_float: yes
  pdf_document: default
  openintro::lab_report: default
editor_options:
  chunk_output_type: console
---

## **The Assignment – Working with XML and JSON in R**

**I picked three movies:**

1- Rich Dad Poor.

2- Shoe Dog: A Memoir by the Creator of Nike.

3- Fear: Trump in the White House.

Other than the Book Title I am using the four other attributes including Author, Genre, YearPublished, and Number of Pages. I created three files to store the information in HTML, XML, and JSON formats, I will run codes to pull the data from my github to compare the data.

**Please find the [Rpubs here](https://rpubs.com/karim7mod/673226)**

**Please find rmd, html, xml, json files on [Github link](https://github.com/akarimhammoud/CUNY-SPS/tree/master/607-Data-Acquisition-and-Management-CUNY-SPS-Fall2020/W7%20-%20Web%20Technologies%3B%20MongoDB)**

## CSV File

I will start by pulling the CSV file to compare the data

```{r}
df_csv <- read.csv("https://raw.githubusercontent.com/akarimhammoud/CUNY-SPS/master/607-Data-Acquisition-and-Management-CUNY-SPS-Fall2020/W7%20-%20Web%20Technologies%3B%20MongoDB/Books.csv")
df_csv
```

## XML File

```{r}
#running the required libraries
library("RCurl", quietly = TRUE)
library("XML", quietly = TRUE)

#getting the url for the xml data
df_xml <-getURL("https://raw.githubusercontent.com/akarimhammoud/CUNY-SPS/master/607-Data-Acquisition-and-Management-CUNY-SPS-Fall2020/W7%20-%20Web%20Technologies%3B%20MongoDB/Books.xml")

#getting the xml data frame for the file
df_xml <- xmlToDataFrame(df_xml)
df_xml
```

## HTML File

```{r include=TRUE, message=FALSE}
#running the required libraries
library("textreadr", quietly = TRUE)
library("rvest", quietly = TRUE)
library("xml2", quietly = TRUE)
```

```{r}
#getting the url for the HTML data
df_html<-read_html('https://raw.githubusercontent.com/akarimhammoud/CUNY-SPS/master/607-Data-Acquisition-and-Management-CUNY-SPS-Fall2020/W7%20-%20Web%20Technologies%3B%20MongoDB/Books.html')

#Running the html table
df_html <- html_table(df_html, dec = ".", fill = TRUE)
df_html
```

## JSON File

```{r}
#running the required libraries
library(jsonlite, quietly = TRUE)

#getting the url for the JSON data
df_json <- read_json("https://raw.githubusercontent.com/akarimhammoud/CUNY-SPS/master/607-Data-Acquisition-and-Management-CUNY-SPS-Fall2020/W7%20-%20Web%20Technologies%3B%20MongoDB/Books.json", simplifyVector = TRUE)

df_json
```

## Conclusion
**The HTML, XML, and JSON files are identical and look the same after running them on Rmarkdown, even with the two authors case it is still identical.**

**Here a is reference on how to create an [HTML tables](https://www.w3schools.com/html/html_tables.asp)**


**Here a is reference on how to create an [XML tables](http://www.w3schools.com/xml/default.asp)**


**Here a is reference on how to create an [JSON tables](https://www.w3schools.com/js/js_json_intro.asp)**




