---
title: "W9 - Working with web APIs"
author: "Karim Hammoud"
date: "`r Sys.Date()`"
output:
  html_document:
    highlight: pygments
    theme: cerulean
    toc: yes
    toc_float: yes
  pdf_document: default
  openintro::lab_report: default
editor_options:
  chunk_output_type: console
---

### Overview

The New York Times web site provides a rich set of APIs, as described here: https://developer.nytimes.com/apis You’ll need to start by signing up for an API key.
Your task is to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and transform it into an R DataFrame.

**You can find [rpubs here]("https://rpubs.com/karim7mod/680762")**

**And [Github here]("")**



### Environment preperation

```{r}
options(stringsAsFactors = FALSE)
library(httr)
library(jsonlite)
library(DT)
```

```{r echo=FALSE}
# setup a hidden key
apiKey = "eSocQwpySQrEosBhHvPcNVxv9BHsSZpf"
```

### Access and convert the data for the movies API.

**I mainly focused on the movie review API from NYT.**

```{r}
# setup the link for the API
url <- "https://api.nytimes.com/svc/movies/v2/reviews/search.json"

#pull the data
nyt <- paste0(url, '?api-key=', apiKey)

nyt1 <- fromJSON(nyt, flatten = TRUE)
```

### Show the table and conclusion for the movie review table.

**As we see below the resutl is table for the movies review**

```{r}
# setup the table
datatable(nyt1$results)
```
